
\begin{itemize}
\item PP is not proprietary and can be used for any purpose. The only
restriction is that suing of the authors for any damage
the code may cause is not allowed.

\item PP runs on a range of UNIX and UNIX-like operating systems, including
 SUNOS, Ultrix, and BSD.  A full list of platforms on which PP is know
 to run is included in the distribution.

\item Current modules include:

\input{highlights}

\item No User Agents (UAs) are included with PP.  However, procedural access
to the MTA is documented, to encourage others to write or to port UAs.
Several existing UAs, such as MH, may be used with PP. 


\item It is expected that a Message Store to be used in conjunction
  with PP (PPMS), and an associated X--Windows User Agent (XUA) will
  be released on beta test in first quarter 92.

\item The core routing of PP \ppversion/ is table based.   DNS is used
  by the SMTP channel.   The next version of PP will support Directory
  Based routing, which may use X.500 or DNS.   

\item PP \ppversion/ requires ISODE \isodevrsn/.

\item X--Windows release X11R4 (or greater) is needed by some of the
management tools.  PP  can be operated without these tools.

\item Although PP is not ``supported'' per se (but see later), it does
have a problem reporting address:  

\begin{tabbing}

RFC 822: \= PP-SUPPORT@CS.UCL.AC.UK \\ \\

X.400: \> S=PP-Support; OU=CS; O=UCL; \\
       \> PRMD=UK.AC; ADMD=\verb*| |; C=GB; \\

\end{tabbing}




Bug reports (and fixes) are welcome.

\item The discussion group PP-PEOPLE@CS.UCL.AC.UK is used as an open forum
on PP; Contact PP-PEOPLE-REQUEST@CS.UCL.AC.UK to be added to this
list.

\item The primary documentation for this release consists of a three and
a half volume User's Manual (approx. 300 pages) and a set of UNIX
manual pages.  The sources to the User's Manual are in LaTeX
format.
\end{itemize}

For more information, contact:
\begin{quote}
Steve Kille \\
Department of Computer Science \\
University College London \\
Gower Street \\
London \\
WC1E 6BT \\
England \\
\end{quote}

