\subsection*{DISTRIBUTION SITES}

\subsubsection*{NIFTP}
If you run NIFTP over the public X.25 or over Janet, and are
registered in the NRS at Salford, you can use NIFTP with username
``guest'' and your own name as password, to access UK.AC.UCL.CS to
retrieve the file \uktarfile/.  The
file \ukcompressfile/ is the \pgm{tar} image after being run through the
compress program (approx \compresssize/).


\subsubsection*{FTP}
If you can FTP to the Internet from outside Europe, then use anonymous
FTP to uu.psi.com [136.161.128.3] to retrieve the file \compressfile/ in
binary mode from the isode/ directory.  This file is the \pgm{tar} image
after being run through the compress program and is approximately
\compresssize/ in size.

\medskip
If you can FTP to the Internet from Europe, then use anonymous FTP to
archive.eu.net [192.16.202.1] to retrieve the file \compressfile/ in binary
mode from the network/isode/ directory. This file is the \pgm{tar} image
after being run through the compress program and is approximately
\compresssize/ in size.


\subsubsection*{EUROPE (tape and documentation)}

For mailings in EUROPE, send a cheque or bankers draft and a purchase
order
for 200 Pounds Sterling to:
\[\begin{tabular}{ll}
Postal address:&        Department of Computer Science\\
&                       Attn: Natalie May/Dawn Bailey\\
&                       University College London\\
&                       Gower Street\\
&                       London, WC1E 6BT\\
&                       UK
\end{tabular}\]
For information only:
\[\begin{tabular}{ll}
\ Telephone:&           +44 71--380--7214\\
\ Fax:&                 +44 71--387--1397\\
\ Telex:&               28722\\
\ Internet:&            \verb"natalie@cs.ucl.ac.uk"\\
&                       \verb"dawn@cs.ucl.ac.uk"
\end{tabular}\]
Specify one:
\begin{enumerate}
\item   1600bpi 1/2--inch tape, or

\item   Sun 1/4--inch cartridge tape.
\end{enumerate}
The tape will be written in \pgm{tar} format and returned with
a documentation set.
Do not send tapes or envelopes.
Documentation only is the same price.

\subsubsection*{EUROPE (tape only)}
Tapes without hardcopy documentation can be obtained via the software
distribution service of the European UNIX User group (EUUG).
The PP \ppversion/ distribution is called ``EUUGD22 - PP (\ppversion/)''.
\[\begin{tabular}{ll}
Postal address:&        EUUG Software Distributions\\
&                       c/o Frank Kuiper\\
&                       Centrum voor Wiskunde en Informatica\\
&                       Kruislaan 413\\
&                       1098 SJ  Amsterdam\\
&                       The Netherlands\\[0.1in]
For information only:&\\
\ Telephone:&           +31 20--5924121\\
        &               (or +31 20--5929333)\\
\ Telex:&               12571 mactr nl\\
\ Telefax:&             +31--20--5924199\\
\ Internet:&            \verb"euug-tapes@cwi.nl"
\end{tabular}\]
Specify one:
\begin{enumerate}
\item   1600bpi 1/2--inch tape: 130 Dutch guilders.

\item   800bpi 1/2--inch tape: 150 Dutch guilders.

\item   Sun 1/4--inch cartridge tape (QIC-24 format): 190 Dutch guilders.

\item   Sun 1/4--inch cartridge tape (QIC-11 format): 215 Dutch guilders.
\end{enumerate}
If you require DHL, this is possible and will be billed through.
Note that if you are not a member of EUUG,
then there is an additional handling fee of 300 Dutch guilders
(please enclose a copy of your membership or contribution payment form when
ordering).
Do not send money, cheques, tapes or envelopes;
you will be invoiced.


\subsubsection*{NORTH AMERICA}
For mailings in NORTH AMERICA,
send a check for 375 US dollars to:
\[\begin{tabular}{ll}
Postal address:&University of Pennsylvania\\
&               \small Department of Computer and Information Science\\
&               Moore School\\
&               Attn: David J. Farber (PP Distribution)\\
&               200 South 33rd Street\\
&               Philadelphia, PA 19104-6314\\
&               U.S.A.\\[0.1in]
Telephone:&     +1 215--898--8560
\end{tabular}\]
Specify one:
\begin{enumerate}
\item   1600bpi 1/2--inch tape, or

\item   Sun 1/4--inch cartridge tape.
\end{enumerate}
The tape will be written in \pgm{tar} format and returned with
a documentation set.
Do not send tapes or envelopes.
Documentation only is the same price.


\subsubsection*{AUSTRALIA and NEW ZEALAND}
For mailings in AUSTRALIA and NEW ZEALAND,

\begin{enumerate}
\item If you already have ISODE \isodevrsn/, send a cheque for 100
        dollars Australian to the following address. You will
        recieve only the PP manuals and a PP tape. {\bf Clearly specify
        on the cheque and order that you only want PP \ppversion/.}
\item	If you do not have ISODE \isodevrsn/, send a cheque for 350
        dollars Australian to the following address. You will
        recieve a tape containing both ISODE \isodevrsn/ and PP
	\ppversion/, together 
        with manuals for both ISODE \isodevrsn/ and PP \ppversion/.
\end{enumerate}

\[\begin{tabular}{ll}
Postal address:&        CSIRO DIT\\
&                       Attn: Andrew Waugh (ISODE Distribution)\\
&                       723 Swanston St\\
&                       Carlton, 3053\\
&                       Australia
\end{tabular}\]
For information only:
\[\begin{tabular}{ll}
\ Telephone:&           +61 3--282--2615\\
\ Fax:&                 +61 3--282--2600\\
\ Internet:&            \verb"ajw@mel.dit.csiro.au"
\end{tabular}\]
Specify one:
\begin{enumerate}
\item   1600/3200/6250bpi 1/2--inch tape, or

\item   Sun 1/4-inch cartridge tape in either QIC-11, QIC-24 or QIC-150 format.
\end{enumerate}
The tape will be written in \pgm{tar} format and returned with a documentation set.
Do not send tapes or envelopes.
Documentation only is the same price.

\subsubsection*{FTAM on Janet or PSS}
The source code is available by FTAM at the University College London over X.25
using Janet (DTE \verb"00000511160013"), PSS (DTE \verb"23421920030013") 
or IXI (DTE \verb"20433450420113") with
Transport Selector~\verb"259" (ASCII encoding).
Use the ``anon'' user-identity and retrieve the file \tarfile/.
The file \compressfile/ is the \pgm{tar} image after being
run through the compress program (\compresssize/).

\subsubsection*{FTAM on the Internet}

The source code is available by FTAM over the
Internet at host \verb"osi.nyser.net" \verb"[192.33.4.10]"
(TCP port~102 selects the OSI transport service)
with Transport Selector~\verb"259" (numeric encoding).
Use the ``anon'' user-identity, supply any password,
and retrieve \compressfile/ from the isode/ directory.
This file is the \pgm{tar} image after being run through the compress program
and is approximately \compresssize/ in size.

\subsubsection*{Online Postscript}

For distributions via FTAM,
the file service is provided by the FTAM implementation in ISODE~5.0 or later
(IS FTAM).

For distributions via either FTAM or FTP, there are additional files
available for retrieval, called \ukpsfile/  and \psfile/
which are compressed \pgm{tar} images (\pssize/) containing the entire
documentation set in PostScript format, 
for A4 and US paper respectively.   

\subsection*{SUPPORT}

A UK company has been set up to provide support for the ISODE, PP,
and associated packages: X-Tel Services Ltd.  This company provides an
update service, general assistance and site specific support.
This company is closely associated with the development of PP.

\[\begin{tabular}{ll}
Postal Address:&	X-Tel Services Ltd.\\
&		Nottingham University\\
&		Nottingham, NG7 2RD\\
&	        UK\\
Telephone:&	+44 602-412648\\
Fax:&		+44 602-790278\\
Internet:&	support@xtel.co.uk\\
\end{tabular}\]
