% -*- LaTeX -*-

\chapter{Introduction}

This manual is concerned with the user-visible tools associated with
PP. Other tools which are management oriented or system related are
described in other volumes.

Most of the work of a message transfer system is not visible to the
user, who typically interacts with some intermediary (user agent or
UA). As such this manual is a collection of the tools and interfaces
that are directly visible to the user.  Conceptually, they may be
considered as a part of the UA, which for convenience and performance
are implemented as a part of the MTA.

The main components of this volume consider the following:

\begin{describe}
\item[\verb|local delivery|:]	Local delivery is one of the more
obviously noticeable parts of PP to the user. This is where the
message leaves the transfer system and enters the user's domain. As
such, this interface is tailorable to a considerable degree. PP does
not restrict the user to a given delivery program, but can be
configured on both a per recipient and a per message-type basis. This
is conceptually a user agent function, but it is often more efficient
to do this operation at the user agent/MTA boundary.

\item[\verb|distribution lists|:]	Maintenance and use of
distribution lists is another area where the end user is directly
involved. This verges on administration, but is administration that
can be delegated to given users for each list. The areas discussed
here are the administration of distribution lists and the information
associated with such lists.

\item[\verb|user oriented management|:]	There are certain
administrative functions that it is useful to make available to end
users. These include checking the progress of individual messages and
checking the validity of addresses and usernames.

\end{describe}

Much of the information presented here is often available in the
online \unix/ manuals but usually in a much less verbose format.
These are listed here:
\begin{describe}
\item[\man chkmf (1)]	Check the \file{.mailfilter} syntax.
\item[\man ckmail (1)]	Check status of your mail.
\item[\man flagmail (1)]	Notify a server program of new mail.
\item[\man mail (1)]	A very dumb mail submission program.
\item[\man malias (1)]	Show what mailing lists expand to.
\item[\man mlist (1)]	Distribution list maintenance program.
\item[\man ttyalert (1)]	Alert of new mail on the terminal.
\item[\man xalert (1)]	Alert of new mail in an X window.
\item[\man resend (1)]	Resubmit a message to more recipients.
\item[\man strip\_addr (1)] Remove comments and extraneous parts from an
RFC~822 \linebreak[3]address.

\item[\man mailfilter(5)] Brief description of the
			\file{.mailfilter} syntax.
\end{describe}
